A First Effort for a Distributed Segment-based Approach on Self-Assembled Nano Networks

### 2014-02-14 PAPER/implementation.tex
DiSR logo

A possible hardware implementation of the DiSR algorithm will be
described in this section.  Before to enter in to detail of  DiSR
architecture,  an idea of node structure tailored for this particular
scenario can be found depicted in Fig.~\ref{fig:node_structure}. In
particular, such node should have the following foundamental elements:

@@

	\item \textbf{Switch Matrix:} Driven by a switch controller, enables the
	reciprocal connections beetween		 devices inside the node. Before
	the creation of segments, such controller receives information from
	DiSR block.  After this phase, the switch controller will be driven by
	the block wich implements the routing algorithm. The Switch matrix is
	essentially a series of multiplexers and demultiplexers.

@@

	\item \textbf{DiSR-Routing:} The DiSR block contains all the hardware, control
	logic and configuration registers, related to the implementation of the
	proposed approach. The routing algorithm receives information
	from DiSR  wich indicates the status of the segments that interests a
	particular node. The routing operation will take in account of these
	informations in order to perform a deadlock-free routing. Both Routing
	and DiSR are connected to the switch matrix in order to receive
	packets from the input buffer. Since the packets are processed one at
	a time, a specific arbiter should be present within the switch matrix
	controller.

@@

\begin{figure}
  \centering
  \includegraphics[width=0.45\textwidth]{pictures/node_structure.eps}
  \caption{A possible node structure tailored for a DNA nanonetwork.}
 \label{fig:node_structure}
\end{figure}
%--------------------------------------------------------------------
\subsection{DiSR Architecture}
\label{ssec:disr_architecture}

\begin{figure}
  \centering
  \includegraphics[width=0.40\textwidth]{pictures/disr_rtl_updated.eps}
  \caption{\emph{DiSR} block architecture.}
 \label{fig:implementation}
\end{figure}

To give a basic estimate of the overhead needed we will focus on DiSR-
specific components, that is, the control logic and configuration
registers needed to implement DiSR. In Figure~\ref{fig:implementation}
is shown a sketch of the  architecture, which mainly consist in the
following building elements:

@@

    \item \textbf{Control circuitry:}
    This circuitry decodes the the      incoming packet, the LED
	registers and the DBS,     updating them when required. This block
	drives the LED register   to store or, to read the LED data
	according with \emph{write enable (WE)} or \emph{read enable (RE)}
	signals. Then, the resulting \emph{Ctrl-Out} drives the other
	communication resources for actuating the DiSR routing operation.

@@

While the elements mentioned above are part of the DiSR circuitry,
the other devices depicted in Fig.~\ref{fig:implementation} like the
multiplexer and demultiplexer are implemented outside DiSR. In
particular, such components will be inside the switch matrix  (refers
to the Fig~\ref{fig:node_structure}).

@@

\begin{figure}
  \centering
  \includegraphics[width=0.40\textwidth]{pictures/pipeline.eps}
  \caption{Pipeline description of DiSR operations}
 \label{fig:pipeline}
\end{figure}

Each of just described building block, reacts with the rising edge  of
the \emph{clock signal}. An asynchronous system \emph{reset} signal is
also present to restore all registers to their default values. When
this signal is set, the DBS state machines for each network
node are set as \emph{Free}.  
In Fig.~\ref{fig:pipeline} is described the pipeline structure of the
DiSR architecture. The packets incoming from the neighbours node, or
locally generated will be send to the DBS state machine in order to te
ensure the status commutation. While the DBS updates its state,  in the
next clock cycle the control circuitry will update the Led status and
will drive the switch matrix in order to send eventually the packets to
the  right output buffer.
%--------------------------------------------------------------------
\subsection{Synthesis results}

As discussed above, one of the main design challenge of DNA Self-
Assembled systems is the limited number of resources available to implement both
computations and routing decisions for each node. Assuming a budget of
$10^4$ CNFETs for each network's node~\cite{liu_jetcs}  we estimated
the required resources to implement the entire DiSR block. The RTL
description of the circuitry described in the last section (with an
hardware description language)  of the DiSR circuitry has been written
and synthesized at gate-level using \emph{Synopsys Design Compiler} with 
a generic technology library (GTECH).
Considering the specific layout of each single logic elements  (NAND,
full-adder, latch etc.), it has been possible to get a rough estimate
of the number of transistors necessary to implement the DiSR logic.
Figure~\ref{fig:synthesys} shows the results of synthesis in terms of
number of devices (CNFETs) versus the number of network nodes while
the network scales up from $10\times10$ to $100\times100$ nodes.

\begin{figure}
  \centering
  \includegraphics[width=0.50\textwidth]{pictures/synthesis.eps}
  \caption{Number of network's node (N) vs number of CNFETs necassary to implement
  the DiSR circuitry.}
 \label{fig:synthesys}
\end{figure}

In particular, observing the Figure~\ref{fig:synthesys} can  be
pointed out that the proposed implementation occupies about 17,5\% of
the node budget. In particular the main contribution is due to the
controller circuitry followed by LED registers. Further, more than the
absolute number of devices itself, it is interesting to observe that
the circuitry complexity which implements the DiSR algorithm increases
with a slowly growing trend.   An intuitive explanation of this is the
relatively simple logic of DiSR, whose behaviour is almost code on
scalable storage structures. For example, the number of registers
implementing the \emph{link\_visited[]} and \emph{link\_tvisited[]}
table follow the logarithmic function $N_{reg}=N_{port} \cdot
log_2(N)$ where \emph{Nport} is the number of the router’s ports, N is
the number of network’s nodes.

In order to have an idea of the maximum working clock frequency for
the implemented circuitry, timing results can be obtained from a gate-
level netlist. Since a complete technology library usefull for
commercial synthesis tool (like Design Compiler) doesn't exist for 
this particular technology (at the best that we know), we have
synthesized the RTL description of DiSR with a standard 32nm CMOS
library obtaining  a delay of about $\mathrm{\tau_{DiSR}=10} FO4$
(fan-out of four). Considering the results obtained
in~\cite{deng_isscc07}, wich reports the ratio in therms of FO4
between a standard 32nm CMOS technology and a carbon nanotube ones, 
rough delay estimation can be obtained. In particolar has been reported a
ratio R of:  
\begin{equation}
\mathrm{R=\frac{FO4_{CMOS}}{FO4_{CNFET}}=2}  
\end{equation}
Considering that F04 CMOS is about 30 ps, and the results of last
equation, a FO4 for the CNFET case is equal to about 15 ps, and
then the working frequency can be computed as:
\begin{equation}
\mathrm{f_{clk}=\frac{1}{T_{max}}=\frac{1}{\tau_{DiSR} \cdot FO4_{CNFET}}}=
\frac{1}{10\cdot 15 \cdot 10^{-12}}=6,6 GHz
\end{equation}
where $\mathrm{\tau_{DiSR}}$ was expressed in therms of FO4.

Regards the power consumption of the entire DiSR devices, 
an accurate power analysis with the switch activity
annotation can be avoided because the DiSR circuitry will be active 
only once at the system startup. Indeed, after the setup phase, when all segments
are mapped in to the network, the DiSR block will stop its activity
giving as output the fixed status of segments to the routing
algorithm. For this reason the dynamic power will going  to zero and
the only cosumption will be eventually only due to the leackage
consumption.

>>>>

A possible hardware implementation of the \disr{} algorithm will be
described in this section.  Before to enter in to detail of  \disr{}
architecture,  an idea of node structure tailored for this particular
scenario can be found depicted in Fig.~\ref{fig:node_structure}. In
particular, such node should have the following foundamental elements:

@@

	\item \textbf{Switch Matrix:} Driven by a switch controller, enables the
	reciprocal connections beetween		 devices inside the node. Before
	the creation of segments, such controller receives information from
	\disr{} block.  After this phase, the switch controller will be driven by
	the block wich implements the routing algorithm. The Switch matrix is
	essentially a series of multiplexers and demultiplexers.

@@

	\item \textbf{\disr{}-Routing:} The \disr{} block contains all the hardware, control
	logic and configuration registers, related to the implementation of the
	proposed approach. The routing algorithm receives information
	from \disr{}  wich indicates the status of the segments that interests a
	particular node. The routing operation will take in account of these
	informations in order to perform a deadlock-free routing. Both Routing
	and \disr{} are connected to the switch matrix in order to receive
	packets from the input buffer. Since the packets are processed one at
	a time, a specific arbiter should be present within the switch matrix
	controller.

@@

\begin{figure}
  \centering
  \includegraphics[width=0.45\textwidth]{pictures/node_structure.eps}
  \caption{A possible node structure tailored for a DNA nanonetwork.}
 \label{fig:node_structure}
\end{figure}
%--------------------------------------------------------------------
\subsection{\disr{} Architecture}
\label{ssec:disr_architecture}

\begin{figure}
  \centering
  \includegraphics[width=0.40\textwidth]{pictures/disr_rtl_updated.eps}
  \caption{\emph{\disr{}} block architecture.}
 \label{fig:implementation}
\end{figure}

To give a basic estimate of the overhead needed we will focus on \disr{}-
specific components, that is, the control logic and configuration
registers needed to implement \disr{}. In Figure~\ref{fig:implementation}
is shown a sketch of the  architecture, which mainly consist in the
following building elements:

@@

    \item \textbf{Control circuitry:}
    This circuitry decodes the the      incoming packet, the LED
	registers and the DBS,     updating them when required. This block
	drives the LED register   to store or, to read the LED data
	according with \emph{write enable (WE)} or \emph{read enable (RE)}
	signals. Then, the resulting \emph{Ctrl-Out} drives the other
	communication resources for actuating the \disr{} routing operation.

@@

While the elements mentioned above are part of the \disr{} circuitry,
the other devices depicted in Fig.~\ref{fig:implementation} like the
multiplexer and demultiplexer are implemented outside \disr{}. In
particular, such components will be inside the switch matrix  (refers
to the Fig~\ref{fig:node_structure}).

@@

\begin{figure}
  \centering
  \includegraphics[width=0.40\textwidth]{pictures/pipeline.eps}
  \caption{Pipeline description of \disr{} operations}
 \label{fig:pipeline}
\end{figure}

Each of just described building block, reacts with the rising edge  of
the \emph{clock signal}. An asynchronous system \emph{reset} signal is
also present to restore all registers to their default values. When
this signal is set, the DBS state machines for each network
node are set as \emph{Free}.  
In Fig.~\ref{fig:pipeline} is described the pipeline structure of the
\disr{} architecture. The packets incoming from the neighbours node, or
locally generated will be send to the DBS state machine in order to te
ensure the status commutation. While the DBS updates its state,  in the
next clock cycle the control circuitry will update the Led status and
will drive the switch matrix in order to send eventually the packets to
the  right output buffer.
%--------------------------------------------------------------------
\subsection{Synthesis results}

As discussed above, one of the main design challenge of DNA Self-
Assembled systems is the limited number of resources available to implement both
computations and routing decisions for each node. Assuming a budget of
$10^4$ CNFETs for each network's node~\cite{liu_jetcs}  we estimated
the required resources to implement the entire \disr{} block. The RTL
description of the circuitry described in the last section (with an
hardware description language)  of the \disr{} circuitry has been written
and synthesized at gate-level using \emph{Synopsys Design Compiler} with 
a generic technology library (GTECH).
Considering the specific layout of each single logic elements  (NAND,
full-adder, latch etc.), it has been possible to get a rough estimate
of the number of transistors necessary to implement the \disr{} logic.
Figure~\ref{fig:synthesys} shows the results of synthesis in terms of
number of devices (CNFETs) versus the number of network nodes while
the network scales up from $10\times10$ to $100\times100$ nodes.

\begin{figure}
  \centering
  \includegraphics[width=0.50\textwidth]{pictures/synthesis.eps}
  \caption{Number of network's node (N) vs number of CNFETs necassary to implement
  the \disr{} circuitry.}
 \label{fig:synthesys}
\end{figure}

In particular, observing the Figure~\ref{fig:synthesys} can  be
pointed out that the proposed implementation occupies about 17,5\% of
the node budget. In particular the main contribution is due to the
controller circuitry followed by LED registers. Further, more than the
absolute number of devices itself, it is interesting to observe that
the circuitry complexity which implements the \disr{} algorithm increases
with a slowly growing trend.   An intuitive explanation of this is the
relatively simple logic of \disr{}, whose behaviour is almost code on
scalable storage structures. For example, the number of registers
implementing the \emph{link\_visited[]} and \emph{link\_tvisited[]}
table follow the logarithmic function $N_{reg}=N_{port} \cdot
log_2(N)$ where \emph{Nport} is the number of the router’s ports, N is
the number of network’s nodes.

In order to have an idea of the maximum working clock frequency for
the implemented circuitry, timing results can be obtained from a gate-
level netlist. Since a complete technology library usefull for
commercial synthesis tool (like Design Compiler) doesn't exist for 
this particular technology (at the best that we know), we have
synthesized the RTL description of \disr{} with a standard 32nm CMOS
library obtaining  a delay of about $\mathrm{\tau_{\disr{}}=10} FO4$
(fan-out of four). Considering the results obtained
in~\cite{deng_isscc07}, wich reports the ratio in therms of FO4
between a standard 32nm CMOS technology and a carbon nanotube ones, 
rough delay estimation can be obtained. In particolar has been reported a
ratio R of:  
\begin{equation}
\mathrm{R=\frac{FO4_{CMOS}}{FO4_{CNFET}}=2}  
\end{equation}
Considering that F04 CMOS is about 30 ps, and the results of last
equation, a FO4 for the CNFET case is equal to about 15 ps, and
then the working frequency can be computed as:
\begin{equation}
\mathrm{f_{clk}=\frac{1}{T_{max}}=\frac{1}{\tau_{\disr{}} \cdot FO4_{CNFET}}}=
\frac{1}{10\cdot 15 \cdot 10^{-12}}=6,6 GHz
\end{equation}
where $\mathrm{\tau_{\disr{}}}$ was expressed in therms of FO4.

Regards the power consumption of the entire \disr{} devices, 
an accurate power analysis with the switch activity
annotation can be avoided because the \disr{} circuitry will be active 
only once at the system startup. Indeed, after the setup phase, when all segments
are mapped in to the network, the \disr{} block will stop its activity
giving as output the fixed status of segments to the routing
algorithm. For this reason the dynamic power will going  to zero and
the only cosumption will be eventually only due to the leackage
consumption.
